The formalisms of Boolean circuits and binary decision diagrams are central in many
communities with limited interaction: circuit complexity, SAT, knowledge
compilation, etc. This document is an attempt to regroup definitions and
conditions on diagrams and circuits to highlights the connections
between these formalisms, and also mention the connections to automata.







\subparagraph*{Related work (classes that we don't cover)}
We also note that a class of circuits intermediate between DNNFs and SDNNFs has been proposed in~\cite{amarilli2017circuit}: this class requires that there is an order on the variables and that, for any $\land$-gate $g$, then $\var(g_1)$ and $\var(g_2)$ for its two inputs $g_1$ and $g_2$ are contained in disjoint intervals of the ordering. A structured circuit is always ordered, taking for order the traversal order on the leaves of a vtree, but the converse is not necessarily true.

Last, we note that there is a notion of $\land$-BDDs that has been proposed, extending BDDs with $\land$-gates. TODO, in particular for how order works there (see Florent?).

And-OBDDs?

Last, we mention that the determinism and decomposability properties have
analogous definitions in the world of \emph{arithmetic circuits}. Intuitively,
an arithmetic circuit is a circuit computing a polynomial over a given field
$\mathbb{F}$: the internal nodes correspond to multiplication or addition, and
the nodes having no inputs are either elements of $\mathbb{F}$ or variables. We
refer to the survey of \cite{shpilka2010arithmetic} for formal definitions. Then,
decomposability means that the polynomial computed is \emph{multilinear}, i.e.,
in every monomial the degree of each variable is at most one, whereas
determinism means that the coefficients of every monomial is simply $1$.
\antoine{TODO: mention connection between zero-suppressed semantics and
arithmetic circuits; the connection between the monomial expressed by an
arithmetic circuit and the set of accepting valuations only work in the
zero-suppressed semantics}
\antoine{TODO: mention a bit the possibility of doing conversions}
\antoine{TODO: also mention depth reduction}

